### 2024-07-02 scripts/explore_analysis.Rmd
updated figures and explore analysis

---
title: "Drought vulnerability of tree growth"
author: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## HKK climate

```{r echo = FALSE, message=F, warning=F}

#load libraries
library(tidyverse)
library(lubridate)
library(httr)
library(janitor)

#reading the GitHub PAT (stored privately and locally)
pat<-as.character(read.table("../data/HKK-dendro/git_pat.txt")[1,1])

# remotes:::download("DendroByBand.RData",
#                    "https://raw.githubusercontent.com/forestgeo/HKK-dendrobands/main/data/modified_data/DendroByBand.RData",
#                    auth_token = pat)

load("../data/HKK-dendro/DendroByBand.RData")

dendrobyband<-merge(dendrobyband, trees, by="Tag")

hkk_OBSID$year<-year(hkk_OBSID$Date2)
hkk_OBSID$month<-month(hkk_OBSID$Date2)
hkk_OBSID$day<-day(hkk_OBSID$Date2)

#correcting two sets of wrong dates

hkk_OBSID$year[hkk_OBSID$year==2209]<-2009
hkk_OBSID$year[hkk_OBSID$month==1 & hkk_OBSID$Cno==5]<-2011

dendrobyband<-merge(dendrobyband, hkk_OBSID, by=c("OBSID", "Cno", "Date2", "Tag"))

dendrobyband$Date2<-as.Date(paste0(dendrobyband$year,"-", dendrobyband$month, "-", dendrobyband$day))

#this can be refined if needed
#wet_census<-c(1, 3, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28)

#getting climate data for all

#CHIRPS data ----

# CHIRPS data GEE code here - https://code.earthengine.google.com/894597bd093906fb0ff82320f29c0842

chirps_hkk<-read.csv("../data/climate/CHIRPS_from_GEE.csv")
colnames(chirps_hkk)[1]<-"time"

#making dates manageable
chirps_hkk<- chirps_hkk %>%
  dplyr::mutate(time=as.Date(time, format = "%b %d, %Y"),
                year=year(time),
                month=month(time))

census_int<-dendrobyband%>%
  group_by(Cno) %>%
  dplyr::summarise(start=as.Date(min(Date2, na.rm = T)),
                   end=as.Date(max(Date2, na.rm=T)))#%>%
  #dplyr::mutate(type=ifelse(Cno %in% wet_census, "wet", "dry"))

spei_hkk<-read.csv("../data/climate/SPEI_HKK_from_GEE.csv")
colnames(spei_hkk)[1]<-"time"

#making dates manageable
spei_hkk<- spei_hkk %>%
  dplyr::mutate(time=as.Date(time, format = "%b %d, %Y"),
                year=year(time),
                month=month(time))

pre_yr_census<-ggplot()+
  #geom_area(data=chirps_hkk, aes(x=time, y=precipitation), fill="blue", alpha=0.5)+
  geom_line(data=chirps_hkk, aes(x=time, y=precipitation), col="blue", alpha=0.5)+
  geom_line(data=spei_hkk, aes(x=time, y=(100*SPEI_01_month-250)), col="red")+
  geom_rect(data=census_int, aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf), 
              fill = "grey10", alpha = 0.5) +
  geom_text(data=census_int, aes(x=start, y=500, 
                                 label= paste0("C", 1:29), angle=90), size=2)+
   scale_y_continuous(    # Features of the first axis
     name = "monthly rainfall (mm)",

     # Add a second axis and specify its features
     sec.axis = sec_axis(~(.+250)*0.01, name="SPEI"))+
  xlab("")+theme_bw()+ ylab("total monthly precipitation (mm) from CHIRPS")+
  ggtitle("HKK dendroband censuses and climate")+
  #scale_x_continuous(n.breaks=15)+
  theme(axis.text.x = element_text(angle=45))

png("../doc/display/census_precip.png", width=6, height=3, res=300, units="in") 
pre_yr_census
dev.off()

#rainfall per year

#-----------------Decisions on climate data-----------------------=
# Summarising climate data to relevant time window for SCBI
# As per issue #42 and KAT pub https://onlinelibrary-wiley-com.smithsonian.idm.oclc.org/doi/full/10.1111/gcb.15934
# HKK best climate variables are:
# PPT - previous Sep to current June (it is actually precipitation day freq, but using sum now)
# Tmax - current Apr to Oct
#---------------------------------------------------------------=

#applying these decisions and rescaling

#mean annual precip from 1981 to 2023 is 1698.4

#code - https://code.earthengine.google.com/b0eda787f2e1ba578bbf50a8a9d04843

climate_data_yr<-chirps_hkk %>%
  group_by(year) %>%
  summarise(
    tot_precip=sum(precipitation),
    pre_SeptoDec=sum(precipitation[month>=9]), #Sep to Dec precip for each year
    pre_JantoJun=sum(precipitation[month<=6]) #Jan to June precip for each year
    #tmx_window=sum(value[variable == "tmx"& month %in% 4:10])
)%>% #sum of potential evapotranspiration for current May to July
  mutate(
    pre_window=pre_JantoJun+lag(pre_SeptoDec), #adding prev year's June to Dec and current year's Jan to June
    tot_pre_diff=tot_precip-1698.4
    )

#standard deviation is 254.06

precip_plot<-ggplot(climate_data_yr)+
  geom_bar(aes(x=year, y=tot_pre_diff), stat="identity")+
  ylab("difference from 40-year mean (mm)")+
  geom_hline(yintercept = c(254.06, -254.06), lty=2)+
  theme_bw()
precip_plot

# library
library(ggridges)
library(ggplot2)
library(viridis)
library(hrbrthemes)

chirps_hkk$year_fac<-as.character(chirps_hkk$year)
chirps_hkk$year_fac<-factor(chirps_hkk$year)

# Plot
ggplot(chirps_hkk, aes(x = precipitation, y = year_fac, 
                       fill = stat(x))) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(name = "rainfall", option = "C") +
  labs(title = 'Rainfall in HKK') +
  #theme_ipsum() +
  theme_bw()+
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )

```
```{r echo=F}

#remove all data with decimal errors, misplaced decimal

dendro.data<-dendrobyband %>%
  dplyr::rename(sp=SPCODE.UPDATE)%>%
  mutate(year=year(Date2)) %>%
  #filter(year!=2209)%>% #quick fix to remove bad dates - there are 30 entries
  #filter(start_date <= year & year <= end_date) %>%
  #filter(Cno %in% wet_census)%>%
  filter(is.na(DBH1_flag), DBH1_error=="none",
         win.decimal.error=="none", is.na(win.flag))%>%
  group_by(Cno, sp, Tag) %>%
  dplyr::summarise(calcDBH=calcDBH[which(n.dendro==min(n.dendro))],
                   Date2=Date2[which(n.dendro==min(n.dendro))],
                   Cii=max(Cii, na.rm=T))

#first fitting a line through the time series
obs.model.dbh<-nlme::lme(calcDBH~Cno, random=~1|Tag, data=dendro.data,
                         na.action=na.omit,
                         method="ML", control=(msMaxIter=100)) #had to increase the number of iterations

#standardised residuals calculated https://stat.ethz.ch/R-manual/R-devel/library/nlme/html/residuals.lme.html

std.res.dbh<-residuals(obs.model.dbh, type="pearson")
hist(std.res.dbh)
dendro.data$std.res<-rep(NA, nrow(dendro.data))
dendro.data$std.res[!is.na(dendro.data$calcDBH)]<-std.res.dbh
#defining all observation with abs(residual) >3 as an outlier
dendro.data$out<-factor(ifelse(abs(dendro.data$std.res)<=3, 0, 1))

#find the trees which have outlier observations
outs.Tag<-unique(dendro.data$Tag[dendro.data$out==1])

#plot trees that have outliers
ggplot(dendro.data[dendro.data$Tag %in%outs.Tag,],
       aes(x = Cno, y = calcDBH, col=factor(out), group = Tag))+
  geom_line(alpha=0.5)+geom_point()+
  scale_color_manual(values=c("grey20", "red", "grey"))

dendro.data.noouts<-dendro.data %>%
  filter(abs(std.res)<=3)%>%
  select(-c(std.res, out))
  #mutate(dbh3=ifelse(abs(std.res)<=3, calcDBH, NA)) #%>% #create an outlier removed column

```
```{r calculating increments for specific years}

dendro_2010<- dendro.data.noouts %>%
  filter(Cno %in% c(1, 3, 5))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(calcDBH_min1=dplyr::lag(calcDBH, n=1, default=NA, order_by=Tag),
         cii_min1=dplyr::lag(Cii, n=1, default=NA, order_by=Tag),
         tdif=difftime(Date2, lag(Date2, n=1, default=NA, order_by=Tag), units = "days"),
         inc=calcDBH-calcDBH_min1,
         inc_annual=(inc/as.numeric(tdif))*365,
         inc.time=Cno-lag(Cno),
         #removing increment value if the previous measure was before the last year
         #inc=ifelse(inc.time==1, inc, NA),
         inc_annual=ifelse(inc.time==2, inc_annual, NA),
         treeID=paste0("HKK.", Tag))%>%
  filter(!is.na(inc))%>%
  filter(!is.na(inc_annual))%>%
  ungroup()%>%
  dplyr::mutate(
    diainc.scaled = scale(inc_annual),
    out.inc = ifelse(abs(diainc.scaled)>3, 1, 0))%>%
  filter(out.inc==0)

quantiles_2009<-quantile(dendro_2010$inc_annual[dendro_2010$Cno==3], 
                         probs=c(0.25, 0.5, 0.75, 1))

quantiles_2010<-quantile(dendro_2010$inc_annual[dendro_2010$Cno==5], 
                         probs=c(0.25, 0.5, 0.75, 1))

dendro_2010_plot<-ggplot(data = dendro_2010, aes(x=inc_annual, group=factor(Cno), fill=factor(Cno), col=factor(Cno))) +
  geom_density(alpha=0.5) +
  scale_color_manual(values=c("forestgreen", "brown"))+
  scale_fill_manual(values=c("forestgreen", "brown"))+
  geom_vline(xintercept=quantiles_2009, col="forestgreen")+
  geom_text(aes(x=0.75, y=1, label="2009"), col="forestgreen", size=3)+
  geom_text(aes(x=0.75, y=1.5, label="2010"), col="brown", size=3)+
  geom_vline(xintercept=quantiles_2010, col="brown")+ 
  xlab("diameter increment (cm)") + ylab("density")+
  guides(col="none", fill="none")+
  theme_bw()

dendro_2015<- dendro.data.noouts %>%
  filter(Cno %in% c(11, 13, 15))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(calcDBH_min1=dplyr::lag(calcDBH, n=1, default=NA, order_by=Tag),
         cii_min1=dplyr::lag(Cii, n=1, default=NA, order_by=Tag),
         tdif=difftime(Date2, lag(Date2, n=1, default=NA, order_by=Tag), units = "days"),
         inc=calcDBH-calcDBH_min1,
         inc_annual=(inc/as.numeric(tdif))*365,
         inc.time=Cno-lag(Cno),
         #removing increment value if the previous measure was before the last year
         #inc=ifelse(inc.time==1, inc, NA),
         inc_annual=ifelse(inc.time==2, inc_annual, NA),
         treeID=paste0("HKK.", Tag))%>%
  filter(!is.na(inc))%>%
  filter(!is.na(inc_annual))%>%
  ungroup()%>%
  dplyr::mutate(
    diainc.scaled = scale(inc_annual),
    out.inc = ifelse(abs(diainc.scaled)>3, 1, 0))%>%
  filter(out.inc==0)

quantiles_2014<-quantile(dendro_2015$inc_annual[dendro_2015$Cno==13], 
                         probs=c(0.25, 0.5, 0.75, 1))

quantiles_2015<-quantile(dendro_2015$inc_annual[dendro_2015$Cno==15], 
                         probs=c(0.25, 0.5, 0.75, 1))

dendro_2015_plot<-ggplot(data = dendro_2015, aes(x=inc_annual, group=factor(Cno), fill=factor(Cno), col=factor(Cno))) +
  geom_density(alpha=0.5) +
  scale_color_manual(values=c("forestgreen", "brown"))+
  scale_fill_manual(values=c("forestgreen", "brown"))+
  geom_vline(xintercept=quantiles_2014, col="forestgreen")+
  geom_text(aes(x=0.75, y=1, label="2014"), col="forestgreen", size=3)+
  geom_text(aes(x=0.75, y=1.5, label="2015"), col="brown", size=3)+
  geom_vline(xintercept=quantiles_2015, col="brown")+ 
  xlab("diameter increment (cm)") + ylab("density")+
  guides(col="none", fill="none")+
  theme_bw()

dendro_2019<- dendro.data.noouts %>%
  filter(Cno %in% c(19, 21, 23))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(calcDBH_min1=dplyr::lag(calcDBH, n=1, default=NA, order_by=Tag),
         cii_min1=dplyr::lag(Cii, n=1, default=NA, order_by=Tag),
         tdif=difftime(Date2, lag(Date2, n=1, default=NA, order_by=Tag), units = "days"),
         inc=calcDBH-calcDBH_min1,
         inc_annual=(inc/as.numeric(tdif))*365,
         inc.time=Cno-lag(Cno),
         #removing increment value if the previous measure was before the last year
         #inc=ifelse(inc.time==1, inc, NA),
         inc_annual=ifelse(inc.time==2, inc_annual, NA),
         treeID=paste0("HKK.", Tag))%>%
  filter(!is.na(inc))%>%
  filter(!is.na(inc_annual))%>%
  ungroup()%>%
  dplyr::mutate(
    diainc.scaled = scale(inc_annual),
    out.inc = ifelse(abs(diainc.scaled)>3, 1, 0))%>%
  filter(out.inc==0)

quantiles_2018<-quantile(dendro_2019$inc_annual[dendro_2019$Cno==21], 
                         probs=c(0.25, 0.5, 0.75, 1))

quantiles_2019<-quantile(dendro_2019$inc_annual[dendro_2019$Cno==23], 
                         probs=c(0.25, 0.5, 0.75, 1))

dendro_2019_plot<-ggplot(data = dendro_2019, aes(x=inc_annual, group=factor(Cno), fill=factor(Cno), col=factor(Cno))) +
  geom_density(alpha=0.5) +
  scale_color_manual(values=c("forestgreen", "brown"))+
  scale_fill_manual(values=c("forestgreen", "brown"))+
  geom_vline(xintercept=quantiles_2018, col="forestgreen")+
  geom_text(aes(x=0.75, y=1, label="2018"), col="forestgreen", size=3)+
  geom_text(aes(x=0.75, y=1.5, label="2019"), col="brown", size=3)+
  geom_vline(xintercept=quantiles_2019, col="brown")+ 
  xlab("diameter increment (cm)") + ylab("density")+
  guides(col="none", fill="none")+
  theme_bw()

library(gridExtra)
png("../doc/display/increments_raw.png", width=8, height=4, units="in", res=300)
grid.arrange(dendro_2010_plot, dendro_2015_plot, dendro_2019_plot, nrow=1)
dev.off()

```

```{r long-term growth rates}

library(ggplot2)

pat<-as.character(read.table("../data/HKK-dendro/git_pat.txt")[1,1])

remotes:::download("../data/HKK-dendro/stem1.RData",
                   "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem1.rdata",
                   auth_token = pat)
remotes:::download("../data/HKK-dendro/stem2.RData",
                   "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem2.rdata",
                   auth_token = pat)
remotes:::download("../data/HKK-dendro/stem3.RData",
                   "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem3.rdata",
                   auth_token = pat)
remotes:::download("../data/HKK-dendro/stem4.RData",
                   "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem4.rdata",
                   auth_token = pat)

load("../data/HKK-dendro/stem1.RData")
load("../data/HKK-dendro/stem2.RData")
load("../data/HKK-dendro/stem3.RData")
load("../data/HKK-dendro/stem4.RData")

hkk_census<-bind_rows(hkk.stem1, hkk.stem2, hkk.stem3, hkk.stem4)

hkk_census <- hkk_census %>%
  filter(status=="A")%>%
  filter(tag %in% dendro.data.noouts$Tag)%>%
  group_by(StemTag)%>%
  dplyr::mutate(dbhmin1 = dplyr::lag(dbh, n=1, default=NA, order_by=StemTag),
                inc_annual = (dbh-dbhmin1)/5)%>%
  filter(!is.na(dbhmin1))%>%
  dplyr::summarise(avg_inc = mean(inc_annual, na.rm=T),
                   sd_inc = sd(inc_annual, na.rm=T))

inc_census<-ggplot(data = hkk_census, aes(x=avg_inc)) +
  geom_density(alpha=0.5) +
  xlab("mean annual increment (cm)")+
  theme_bw()

png("../doc/display/census_increment.png", width=4, height=4, units="in", res=300)
inc_census
dev.off()

```

```{r drought resistance}

dec_williams<-read.csv("../data/HKK-dendro/species.csv")

dec_williams<-dec_williams %>%
  filter(n.tree >= 5)

#2010 drought--------

resistance_2010 <- dendro_2010 %>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/inc_min1,
         resistance.dif = inc_annual-inc_min1,
         resistance.prop = (inc_annual-inc_min1)/inc_min1)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

resistance_2010<-merge(resistance_2010, select(dec_williams, -"X"), by="sp", all.x=T)

resistance_2010_div<-ggplot(data = resistance_2010, aes(x=resistance.div)) +
  geom_density(alpha=0.5) +
  theme_bw()

resistance_2010_dif<-ggplot(data = resistance_2010, aes(x=resistance.dif)) +
  geom_density(alpha=0.5) +
  theme_bw()

#2015 drought---------

resistance_2015 <- dendro_2015 %>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/inc_min1,
         resistance.dif = inc_annual-inc_min1,
         resistance.prop = (inc_annual-inc_min1)/inc_min1)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

resistance_2015<-merge(resistance_2015, select(dec_williams, -"X"), by="sp", all.x=T)

resistance_2015_div<-ggplot(data = resistance_2015, aes(x=resistance.div)) +
  geom_density(alpha=0.5) +
  theme_bw()

resistance_2015_dif<-ggplot(data = resistance_2015, aes(x=resistance.dif)) +
  geom_density(alpha=0.5) +
  theme_bw()

#2019 drought--------------

resistance_2019 <- dendro_2019 %>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/inc_min1,
         resistance.dif = inc_annual-inc_min1,
         resistance.prop = (inc_annual-inc_min1)/inc_min1)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

resistance_2019<-merge(resistance_2019, select(dec_williams, -"X"), by="sp", all.x=T)

resistance_2019_div<-ggplot(data = resistance_2019, aes(x=resistance.div)) +
  geom_density(alpha=0.5) +
  theme_bw()

resistance_2019_dif<-ggplot(data = resistance_2019, aes(x=resistance.dif)) +
  geom_density(alpha=0.5) +
  theme_bw()

png("../doc/display/resistance_dif.png", width=4, height=8, units="in", res=300)
grid.arrange(resistance_2010_dif, resistance_2015_dif, resistance_2019_dif, nrow=3)
dev.off()

```

```{r resistance map}

quantiles_2010_dif<-quantile(resistance_2010$resistance.dif, 
                         probs=c(0.25, 0.5, 0.75, 1))

quantiles_2015_dif<-quantile(resistance_2015$resistance.dif, 
                         probs=c(0.25, 0.5, 0.75, 1))

quantiles_2019_dif<-quantile(resistance_2019$resistance.dif, 
                         probs=c(0.25, 0.5, 0.75, 1))

resistance_2010$quantile<-ifelse(resistance_2010$resistance.dif<quantiles_2010_dif[1], 1, ifelse(resistance_2010$resistance.dif<quantiles_2010_dif[2], 2,
            ifelse(resistance_2010$resistance.dif<quantiles_2010_dif[3], 3, 4))

                                 )

resistance_2015$quantile<-ifelse(resistance_2015$resistance.dif<quantiles_2015_dif[1], 1, ifelse(resistance_2015$resistance.dif<quantiles_2015_dif[2], 2,
            ifelse(resistance_2015$resistance.dif<quantiles_2015_dif[3], 3, 4))

                                 )

resistance_2019$quantile<-ifelse(resistance_2019$resistance.dif<quantiles_2019_dif[1], 1, ifelse(resistance_2019$resistance.dif<quantiles_2019_dif[2], 2,
            ifelse(resistance_2019$resistance.dif<quantiles_2019_dif[3], 3, 4))

                                 )

resistance_all<-rbind(resistance_2010, resistance_2015, resistance_2019)
resistance_all$yr<-c(rep(2010, nrow(resistance_2010)), rep(2015, nrow(resistance_2015)), rep(2019, nrow(resistance_2019)))

resistance_all<-merge(resistance_all, select(trees, -"SPCODE.UPDATE"), by="Tag")

resistance_map<-ggplot(resistance_all, aes(x=X, y=Y, size=as.numeric(DBH.04), col=quantile))+
  geom_point(alpha=0.5)+
   scale_size(range = c(.1, 10), name="DBH (cm)")+
  guides(color=guide_legend(title="resistance"))+
  scale_colour_gradient(low="red", high="blue") +
  facet_grid(yr~.)+
  theme_minimal()+
  coord_fixed()

png("../doc/display/resistance_map.png", width=8, height=10, units="in", res=300)
resistance_map
dev.off()
```
```{r basic model}

model_2010<-glm(resistance.dif ~ calcDBH_min1 + cii_min1 + williams_dec+
                  calcDBH_min1*cii_min1 + 
                  calcDBH_min1*williams_dec+
                  cii_min1*williams_dec,
                data=resistance_2010)

summary(model_2010)

model_2015<-glm(resistance.dif ~ calcDBH_min1 + cii_min1 + williams_dec+
                  calcDBH_min1*cii_min1 + 
                  calcDBH_min1*williams_dec+
                  cii_min1*williams_dec,
                data=resistance_2015)

summary(model_2015)

model_2019<-glm(resistance.dif ~ calcDBH_min1 + cii_min1 + williams_dec+
                  calcDBH_min1*cii_min1 + 
                  calcDBH_min1*williams_dec+
                  cii_min1*williams_dec,
                data=resistance_2019)

summary(model_2019)

resistance_all<-rbind(resistance_2010, resistance_2015, resistance_2019)
resistance_all$yr<-c(rep(2010, nrow(resistance_2010)), rep(2015, nrow(resistance_2015)), rep(2019, nrow(resistance_2019)))

resistance_model<-resistance_all %>%
  select(resistance.dif, calcDBH_min1, cii_min1, williams_dec, yr, quantile)

resistance_model<-resistance_model[complete.cases(resistance_model),]

library(lme4)
library(nlme)
model_all<-lme(resistance.dif ~ calcDBH_min1 + cii_min1 + williams_dec+
                  calcDBH_min1*cii_min1 + 
                  calcDBH_min1*williams_dec+
                  cii_min1*williams_dec,
                random=~1|yr,
                data=resistance_model)

model_all
anova(model_all)

model_all_cat<-lme(quantile ~ calcDBH_min1 + cii_min1 + williams_dec+
                  calcDBH_min1*cii_min1 + 
                  calcDBH_min1*williams_dec+
                  cii_min1*williams_dec,
                random=~1|yr,
                data=resistance_model)

model_all_cat
anova(model_all_cat)
```

```{r predictions, echo=F}

library(sjPlot)

plot_model(model_all, type="pred", terms=c("calcDBH_min1","williams_dec"), show.data = T)
plot_model(model_all, type="pred", terms=c("cii_min1"), show.data = T)
plot_model(model_all, type="pred", terms=c("williams_dec"), show.data = T)
plot_model(model_all, type="pred", terms=c("calcDBH_min1"), show.data = T)

plot_model(model_all, type = "est")+
  theme_bw()+ylim(c(-0.25, 0.25))+
  geom_hline(yintercept = 0)

plot_model(model_all_cat, type = "est")+
  theme_bw()+ylim(c(-0.25, 0.25))+
  geom_hline(yintercept = 0)

plot_model(model_all_cat, type="pred", terms=c("calcDBH_min1","williams_dec"), show.data = T)

# Create a new data frame for predictions
new_data <- expand.grid(
  calcDBH_min1 = seq(min(resistance_model$calcDBH_min1), max(resistance_model$calcDBH_min1), length.out = 100),
  cii_min1 = seq(min(resistance_model$cii_min1), max(resistance_model$cii_min1), length.out = 100),
  williams_dec = seq(min(resistance_model$williams_dec), max(resistance_model$williams_dec), length.out = 100)
)

# Generate predictions
new_data$predicted <- predict(model_all, newdata = new_data, level = 0)

# Create the plot
ggplot(resistance_model, aes(x = williams_dec, y = resistance.dif)) +
  geom_point() +
  geom_line(data = new_data[new_data$cii_min1==2,], aes(y = predicted)) +
  theme_minimal() +
  labs(x = "Deciduousness", y = "Drought resistance", title = "Predictions from LME Model")

```

##Distribution of data with size, exposure and deciduousness

```{r echo=F}

#2014 data distribution with size, exposure and deciduousness

png("../doc/display/dbh_cii.png", width=4, height=4, units="in", res=300)
ggplot(growth_2014, aes(x = dbh, y = factor(cii), 
                       fill = stat(x))) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(name = "growth", option = "C", direction=-1) +
  labs(title = 'DBH in 2014 (cm)', y="canopy illumination") +
  #theme_ipsum() +
  theme_bw()+
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )
dev.off()

length(unique(growth_2014$sp))

png("../doc/display/growth_cii.png", width=4, height=4, units="in", res=300)
ggplot(growth_2014, aes(x = diainc, y = factor(cii), 
                       fill = stat(x))) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(name = "growth", option = "C", direction=-1) +
  labs(title = 'growth in 2014 (cm)', y="canopy illumination") +
  #theme_ipsum() +
  theme_bw()+
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )
dev.off()

dec_williams<-read.csv("../data/HKK-dendro/species.csv")

dec_williams<-dec_williams %>%
  filter(n.tree >= 5)

trial<-merge(growth_2014, dec_williams, by="sp")

png("../doc/display/dbh_cii.png", width=4, height=4, units="in", res=300)
ggplot(trial, aes(x = williams_dec, y = factor(Cii), 
                       fill = stat(x))) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(name = "growth", option = "C", direction=-1) +
  labs(title = 'deciduousness', y="canopy illumination") +
  #theme_ipsum() +
  theme_bw()+
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )
dev.off()

```