### 2025-01-16 results/results.qmd
results presentation

---
title: "HKK drought sensitivity results"
author: "Krishna Anujan"
date: "2025-01-10"
format: 
  revealjs:
    theme: simple
    transition: slide
    slideNumber: true
    center: true
    embed-resources: true
---

## Sensitivity of tree growth to precipitation in seasonally dry tropical forests using long-term dendrometer band measurements

## Objective

::: {.columns}
::: {.column width="40%"}

Disentangling the effects of size, exposure and water availability on drought sensitivity of tree growth

:::

::: {.column width="60%"}

![](../doc/display/hypotheses.png)

:::
:::

## Defining drought sensitivity

$sensitivity_{tree} = \frac{drought.year.increment - mean.annual.increment_{tree}}{mean.annual.increment_{tree}}$

![](../doc/display/sens_definition.png)

## Data

::: {.columns}
::: {.column width="60%"}
| **Variable**   | **Description**                                      |
|----------------|------------------------------------------------------|
| <span style="background-color: #000000; color: #ffffff;"> sensitivity</span>    | Late wet/early dry dendrometer band measurements |
| <span style="background-color: #f7b731; color: #ffffff;">exposure   </span>    | Crown Illumination index at previous dendroband census |
| <span style="background-color: #e15f41; color: #ffffff;">size   </span>        | DBH at previous dendroband census |
| <span style="background-color: #546de5; color: #ffffff;">water</span>          | Topographic wetness index |
:::

::: {.column width="40%"}
![](../doc/display/hypotheses.png)
:::

:::

## ENSO drought years were associated with growth declines across species...

::: {.columns}
::: {.column width="50%}
![](../doc/display/spaghetti_top10_new.png)
:::

::: {.column width="50%"}
![](../results/plots/climate_data_timeseries_2007_2019.png)
:::
:::

## ...but we observed large variation in drought sensitivity within and among species

::: {.columns}
::: {.column width="50%"}
![](../doc/display/sens_all.png)
:::

::: {.column width="50%"}
![](../doc/display/sens_top10.png)
:::
:::

##

::: {.columns}
::: {.column width="50%"}
How well do species traits predict average sensitivity of species?

$sensitivity_{tree}~ 1+ (1|Species)$

![](../results/plots/non_negative/ranefs_intercept_bysp_new.png)
:::

::: {.column width="50%"}
![](../results/plots/non_negative/sp_intercept_plot.png)
:::
:::

## Models

We ran 3 different models

1. varying intercepts by species  
$sensitivity_{tree} = \beta_{DBH}DBH_{0} + \beta_{TWI}TWI + \beta_{CII}CII_{0} + (1|Species)$

2. varying intercepts + predictors scaled within species  
$sensitivity_{tree} = \beta_{DBH}DBH_sp_{0} + \beta_{TWI}TWI_sp + \beta_{CII}CII_sp_{0} + (1|Species)$

3. varying slopes + predictors scaled within species  
$sensitivity_{tree} = \beta_{DBH}DBH_sp_{0} + \beta_{TWI}TWI_sp + \beta_{CII}CII_sp_{0} + (1+DBH+CII+TWI|Species)$

##

::: {.columns}
::: {.column width="30%"}
Variation in sensitivity among individuals was explained more by crown exposure than size per se
:::

::: {.column width="70%"}
![](../results/plots/non_negative/coefs_all_plot.png)
:::
:::

## Higher exposure was associated with more negative growth sensitivities across species

##

::: {.columns}
::: {.column width="30%"}
To corroborate these results we modelled growth occurrence as a binary response.

And found similar, stronger influence of exposure on growth occurrence
:::

::: {.column width="70%"}
![](../results/plots/zero_growth/coefs_all_plot.png)
:::
:::

## When modelled without exposure, the effect of size was more negative as exposure is strongly predicted by size

::: {.columns}
::: {.column width="30%"}
![](../doc/display/hypotheses.png)
:::

::: {.column width="70%"}
![](../results/plots/non_negative/coefs_mediation.png)
:::
:::

This implies mediation effects of size on exposure

## Take-home messages

1. In 2015, the stronger ENSO drought, deciduous and slow-growing species had stronger negative sensitivities.
2. While two droughts decreased growth on most trees, more exposed trees were more negatively impacted than their less exposed counterparts
3. While exposure effects were stronger than size effects on drought sensitivity, we see mediation effects.